Supplementary Information

### 2020-11-17 doc/manuscript/MEE_manuscript_SI.Rmd
update site order in Table S1

We also conducted a formal comparison using identical tree-ring and climate data for four species: PSME (Cedar Breaks, Utah), ABAL (Zofin), PIMA (Scotty Creek), and LITU (SCBI; Figs. `r all_traditional_comparisons`). These species were selected for analysis because they have been well-studied in the past, and there was no manipulation of study species based on analysis results.

@@

The ring-width series from each core was standardized via ARSTAN using a 2/3rds $n$ spline, where $n$ is the number of years in the series (*Cook, 1985; Cook & Kairiukstis, 1990- citations in Helcoski*). 
*(The following italic text is plagarized from Helcoski and needs to be reworded:)* *The influence of outliers in all series was reduced using the adaptive power transformation, which also stabilises the variance over time (Cook & Peters, 1997). Next, each series was stabilised using either the average correlation between raw ring-width series (rbar) method or a 1/ 3rds spline method to adjust changes in variance as series replication decreased towards the earlier portion of each chronology (Jones et al., 1997). The 1/3rds spline method was chosen when replication in the inner portion of each chronology (c. the inner 30–50 yr of each record depending on full chronology length) dropped below three trees. Once that step was complete, a robust biweight mean chronology for each species was calculated from the ring-width indices (Cook, 1985). We chose to use residual chronologies because the autoregressive standardisation process in creating them removes much of the tree-level autocorrelation in growth and these chronologies would most likely contain the most conservative information on drivers of interannual growth (Cook, 1985).*

>>>>

We also conducted a formal comparison using identical tree-ring and climate data for four species: PSME (Cedar Breaks, Utah), ABAL ($\v{Z}$of$\'{i}$n), PIMA (Scotty Creek), and LITU (SCBI; Figs. `r all_traditional_comparisons`). These species were selected for analysis because they have been well-studied in the past, and there was no manipulation of study species based on analysis results.

@@

The ring-width series from each core was standardized via ARSTAN using a 2/3rds $n$ spline, where $n$ is the number of years in the series (*Cook, 1985; Cook & Kairiukstis, 1990- citations in Helcoski*). 
*(The following italic text is self-plagarized from Helcoski and needs to be reworded:)* *The influence of outliers in all series was reduced using the adaptive power transformation, which also stabilises the variance over time (Cook & Peters, 1997). Next, each series was stabilised using either the average correlation between raw ring-width series (rbar) method or a 1/ 3rds spline method to adjust changes in variance as series replication decreased towards the earlier portion of each chronology (Jones et al., 1997). The 1/3rds spline method was chosen when replication in the inner portion of each chronology (c. the inner 30–50 yr of each record depending on full chronology length) dropped below three trees. Once that step was complete, a robust biweight mean chronology for each species was calculated from the ring-width indices (Cook, 1985). We chose to use residual chronologies because the autoregressive standardisation process in creating them removes much of the tree-level autocorrelation in growth and these chronologies would most likely contain the most conservative information on drivers of interannual growth (Cook, 1985).*
